\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\title{Real Analysis Syllabus}
\author{}
\date{}
\begin{document}
\maketitle
\section*{Course Description}
This course will cover the basic concepts of real analysis, including limits, 
continuity, differentiability, and integrability. We will also study some important 
examples and applications of these concepts.
\section*{Objectives}
By the end of this course, students should be able to:
\begin{itemize}
\item Understand the basic concepts of real analysis
\item Apply limits, continuity, differentiability, and integrability in various contexts
\item Work with important examples and applications of these concepts
\end{itemize}
\section*{Outline}
\begin{enumerate}
\item Introduction to Limits
\item Continuity and Differentiability
\item Integrability
\item Examples and Applications
\end{enumerate}
\section*{Homework Problems}
Here are some example homework problems for each class:
\subsection*{Class 1: Introduction to Limits}
Problem 1: Show that the limit of a constant function is equal to the constant.
Problem 2: Find the limit of the function f(x) = x^2 as x approaches 0.
\subsection*{Class 2: Continuity and Differentiability}
Problem 3: Prove that a continuous function on an interval is differentiable at every 
interior point of the interval.
Problem 4: Find the derivative of the function f(x) = sin(x) + cos(x).
\subsection*{Class 3: Integrability}
Problem 5: Show that a Riemann integrable function is bounded on its domain.
Problem 6: Compute the integral of the function f(x) = x^2 from 0 to 1.
\subsection*{Class 4: Examples and Applications}
Problem 7: Find the maximum value of the function f(x) = x - sin(x).
Problem 8: Prove that a differentiable function is continuous at every point in its 
domain.
\end{document}

